---
title: "Homework 2"
author: "Kylie Wheelock Riley"
date: "10/4/2019"
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Problem 1

## Importing and tidying Mr. Trash Wheel data
```{r trash wheel import, message = FALSE}
library(readxl)
trash_wheel = read_excel(path = "./data/mr_trash_wheel_data_aug19.xlsx", 1) %>%
  janitor::clean_names() %>% 
  ## creating reasonable variable names
  rename(vol_y3 = volume_cubic_yards, cigs = cigarette_butts) %>% 
  ## omit non-data entries
  select(-x15, -x16, -x17) %>% 
  ## omit rows that do not include dumpster-sepcific data
  drop_na(dumpster:date) %>%
  ## round number of sports balls to the closest integer and convert to integer
  mutate(sports_balls = round(sports_balls, 0)) %>% 
  mutate(sports_balls = as.integer(sports_balls))
```

## Importing and tidying 2017 precipitation data
```{r cleaning 2017 precipitation}
precip_2017 = read_excel(path = "./data/mr_trash_wheel_data_aug19.xlsx", 6, skip = 1) %>% 
  janitor::clean_names() %>% 
  rename(total_in = total) %>% 
  mutate(
    total_in = na_if(total_in, 0),
    year = "2017"
    ) %>% 
  drop_na(total_in, month) %>% 
  select(year, month, total_in)
```

## Importing and tidying 2018 precipitation data
```{r cleaning 2018 precipitation}
precip_2018 = read_excel(path = "./data/mr_trash_wheel_data_aug19.xlsx", 5, skip = 1) %>% 
  janitor::clean_names() %>% 
  rename(total_in = total) %>% 
  mutate(
    total_in = na_if(total_in, 0),
    year = "2018"
    ) %>% 
  drop_na(total_in, month) %>% 
  select(year, month, total_in)
```

## Joining 2017 and 2018 precipitation data
```{r binding 2017 & 2018 precip data}
precip_bind = bind_rows(precip_2017, precip_2018) %>% 
  mutate(month = month.name[month])
    
```

```{r data summary work, include = FALSE}
mean_trash_wheel = mean(pull(trash_wheel, weight_tons))
mean_cubic_yars = mean(pull(trash_wheel, vol_y3))
total_precip_17 = sum(pull(precip_2017, total_in))
total_precip_18 = sum(pull(precip_2018, total_in))
median_balls_17 = median(pull(trash_wheel, sports_balls), year = 2017)
median_balls_18 = median(pull(trash_wheel, sports_balls), year = 2018)
avg_balls_total = mean(pull(trash_wheel, sports_balls))
```

Mr. Trash Wheel is a trash interceptor that removes trash as it enters Baltimore's Inner Harbor. Between May 2014 and June 2019 Mr.Trash Wheel brought in a  total of `r nrow(trash_wheel)` dumpsters full of trash. The average dumpster weighed  `r round(mean_trash_wheel, digits = 2)` tons with a volume of `r round(mean_cubic_yars, digits = 2)` cubic yards. Trash collected in the dumpsters included cigarette butts, plastic bottles, sports balls, and grocery bags, among others. Mr. Trash Wheel is most active just after a precipitation event.  Between 2017 and 2018 there was a total of `r nrow(precip_bind)` months with precipitation. 2017 had lower total precipitation with  `r round(total_precip_17, digits = 2)` inches compared to a total of `r round(total_precip_18, digits = 2)` inches in 2018. The median number of sports balls collected per dumpster in 2017 was `r median_balls_17` and the 2018 median was also `r median_balls_18`.


# Problem 2

## Importing and tidying the monthly poll data
```{r importing and tidying pols_month data, results = "hide"}
## Importing dataset
pols_month = read_csv(file = "./data/pols_month.csv") %>% 
  ## separating month from 1 column into 3
  separate(mon, into = c("year", "month", "day"), sep = "-") %>% 
   mutate(
    ## mutate month into numeric in order to use the month.name var
    month = as.numeric(month),
    month = month.name[month],
    ## converting year to integer to keep consistent in merge
    year = as.integer(year),
    ##create president variable
    president = ifelse(prez_gop > 0, "gop", "dem")
    ) %>%
  select(-prez_dem, -prez_gop, -day) %>% 
  arrange(year, month, desc(year))

```

## Importing and tidying the S&P data
```{r importing and tidying snp data, results = "hide"}
snp_data = read_csv(file = "./data/snp.csv") %>% 
  mutate( 
    date = lubridate::mdy(date)
    ) %>% 
  separate(date, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(
    month = as.numeric(month),
    month = month.name[month], 
    year = as.integer(year),
  ) %>% 
  select(-day) %>% 
  arrange(year, month, desc(year))
  
```

## Importing and tidying unemployment dataset
```{r importing and tidying unelmployment data, results = "hide"}
unemploy_data = read_csv(file = "./data/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month", 
    values_to = "unemploy_pct"
    ) %>%
  mutate(
    year = as.integer(year),
    month = recode(month, "jan" = "January", "feb" = "February", "mar" = "March", "apr" = "April", "may" = "May", "jun" = "June", "jul" = "July", "aug" = "August", "sep" = "September", "oct" = "October", "nov" = "November", "dec" = "December")
  ) %>% 
  drop_na(unemploy_pct) %>%
  arrange(year, month, desc(year))

```


## Merging 
```{r merging pols snp and unemploy data}
## snp_data into pols_month data 
pols_month_snp_data = 
  left_join(pols_month, snp_data, by = c("year", "month"))

## unemploy_data into pols_month and snp combined
pols_snp_unemploy_data =
  left_join(pols_month_snp_data, unemploy_data, by = c("year", "month"))
```

```{r code for summary, include = FALSE}

mean_close = mean(pull(snp_data, close))
snp_max = max(pull(snp_data, close))
snp_min = min(pull(snp_data, close))

```


The data in Problem 2 summarizes the seats held by each U.S political party, the S&P stock growth, and the U.S. unemployment rate over time. 

The monthly poll data contains the politicial parties of the house, senate, and president for each month from `r min(pull(pols_month, year))` to `r max(pull(pols_month, year))` for a total of `r nrow(pols_month)` months of data. 

The S&P stock dataset shows the average S&P close was `r round(mean_close, digits = 2)` points with a high of `r round(snp_max, digits = 2)` in May 2015 and a low of `r round(snp_min, digits = 2)` in January 1950, the last and first months of the data, respectively.

The unemployment dataset shows the unemployment rate of the U.S per month. The maximum unemployment rate was `r max(pull(unemploy_data, unemploy_pct))` percent and the minimum unemployment rate was `r min(pull(unemploy_data, unemploy_pct))`. For 2010, the middle of the last recession, the average unemployment rate was `r mean(pull(unemploy_data, unemploy_pct), year = 2010)` percent. 


# Problem 3

## Importing and tidying baby names data
```{r importing and tidying baby_names data, results = "hide"}
baby_names = read_csv(file = "./data/popular_baby_names.csv") %>% 
  janitor::clean_names() %>%
  rename(year = year_of_birth , name = childs_first_name) %>% 
  mutate(
    year = as.factor(year),
    rank = as.numeric(rank),
    ethnicity = recode(ethnicity,"ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER", "BLACK NON HISP" = "BLACK NON HISPANIC", "WHITE NON HISP" = "WHITE NON HISPANIC"),
    name = as.factor(toupper(name))
    ) %>% 
  distinct() %>% 
  select(year, gender, ethnicity, name, rank, count)
  

```


## Rank of Olivia as a baby name in NYC over time
```{r olivia table, echo = FALSE}
olivia_table = baby_names %>% 
  select(-count) %>% 
  filter(name == "OLIVIA") %>%
  pivot_wider(
  names_from = "year", 
  values_from = "rank"
  ) 
  knitr::kable(olivia_table)

```

 
## Most popular baby boy names in NYC over time
```{r popular boy names table, echo = FALSE}
boy_popular = baby_names %>% 
  filter(rank == 1, gender == "MALE") %>%
  select(-count) %>% 
  pivot_wider(
  names_from = "year", 
  values_from = "name"
  ) 
  knitr::kable(boy_popular)
```


## Scatterplot of rank and count of popular baby names for Non Hispanic White males born in 2016 in NYC
```{r scatterplot, echo = FALSE}
baby_names %>% 
  filter(gender == "MALE", ethnicity == "WHITE NON HISPANIC", year == 2016) %>% 
  ggplot(aes(x = rank, y = count)) + geom_point()

```

